Type Ia Supernova Explosions from Hybrid Carbon-Oxygen-Neon White Dwarf Progenitors

### 2019-02-11 paper/ms.tex
More on Carlyn's paper

\begin{abstract}
The existence of "hybrid" white dwarfs, made of a C/O
core within a O/Ne shell is accepted, and studies indicate that these are 
viable progenitors for thermonuclear (type Ia) supernovae.
Recent work found that the C/O core is mixed with the surrounding O/Ne
as the white dwarf cools prior to accretion, which results in lower central 
C fractions in the massive progenitor than previously assumed. To further 
investigate the efficacy of hybrid white dwarfs as progenitors of 
thermonuclear supernovae, we performed simulations of thermonuclear 
supernovae from a new series of hybrid progenitors that include the 
effects of cooling. The progenitor white dwarf model
was constructed with the one-dimensional stellar evolution code MESA and
represented a star evolved through the phase of unstable interior mixing 
followed by accretion until it reached conditions for the ignition of 
carbon burning. This MESA model was then mapped to a two-dimensional initial 
condition and an explosion simulated with FLASH. For comparison, a similar
simulation of an explosion was performed from a traditional C/O progenitor
white dwarf. By comparing the yields of the explosions, we found that as
with earlier studies, these hybrid models are viable supernova progenitors.
We contrast explosions from these models with explosions 
from previous hybrid models 
{\color{red} without the mixing during the white dwarf cooling}
and traditional C/O progenitors.
\end{abstract}

@@

Supernovae are classified observationally 
by their light curves and spectra~\citep{minkowski41,bertola64,porterfilippenko87,
wheelerharkness1990conf,Fili97}, with the type Ia designation following from
the absence of H in the spectrum and the presence of a specific Si 
line~\citep{filippenko:optical,hillebrandt.niemeyer:type}. These events
have been associated with C burning under degenerate conditions
for some time~\citep{hoylefowler60,arnett.truran.ea:nucleosynthesis},
but discerning the setting(s) of these events is proving difficult
and remains the subject of active research. At present there are three
widely-accpeted scenarios: the {\em single degenerate} scenario,
the {\em double detonation} or {\em sub-Chandrasekhar} scenario, 
the {\em white dwarf merger} or {\em double degenerate} scenario.
We briefly describe these in the subsection that follows.
Also see~\citet{hillebrandt.niemeyer:type,howell2011,hillebrandtetal2013,calderetal2013,roepkesim2018}
for additional discussion.

@@

The single degenerate picture posits a white dwarf gaining mass
from a main-sequence companion. The process relies on a long
period of accretion combined with either steady burning or a 
series of nova explosions to allow the WD to gain at least
$\sim 0.4$ \Msun\ as needed for it to approach the
limiting Chandrasekhar mass~\citep{starrfieldetal2012}. As it approaches
the Chandrasekhar limit, conditions in the compressed core are right
to ignite the thermonuclear buring that will incinerate the star.

Within this progenitor setting, many mechanisms for the explosion have 
been studied with the result being that the outcome is sensitive to
the ignition conditions and the nature of the thermonuclear buring.
Studies indicate that neither a pure deflagration (subsonic flame front)
nor a pure detonation (supersonic burning front) produce ejecta conisitent
with observations \citep{arnett69,roepkeetal07}.  Instead, the models 
that best reproduce the observed stratified ejecta are those in which 
the burning begins as a subsonic deflagration, which allows the star 
to react and expand thus lowering the density, and is then followed by a 
supersonic detonation that incinerates the 
star~\citep{Nomo84,Khokhlov1991Delayed-detonat,HoefKhok96,GameKhokOran05}.
The ``classic'' delayed detonation model is that of 
\citet{Khokhlov1991Delayed-detonat} \citep[See also][]{hoflich.khokhlov.ea:delayed,GameKhokOran05},
and many variations have been 
explored \citep[and references therein]{hillebrandtetal2013,calderetal2013}.
The approach we employ for this study is a variation, the 
deflagration-to-detonation transition (DDT), in which the transition
occurs at the top of a rising Rayleigh-Taylor~\citep{taylor+50,chandra+81} 
unstable plume of hot, burned 
material when it reaches a threshold density.
We describe this approach and our methodology in detail below.

The double detionation picture is a variation of the single degenerate picture.
In this case, a white dwarf accretes material from a companion, but rather
than gaining enough mass to approach the Chandrasekhar limit, a detonation
occurs in the accreted layer that subsequently triggers a detonation
in the underlying white dwarf \citet{woosleyweavertaam80,taam80a,taam80b,
nomoto80,nomoto82b}. Early studies indicated that a detonation in an 
accreted He
layer could produce an inwardly propagating
shock that would ignite a detonation in the C-O core and found
that this scenario could work for a wide range of white dwarf
masses, not just the near-Chandrasekhar case \citep{livne90}, hence the
moniker ``sub-Chandrasekhar'' \citep{ww94}. Multidimensional supported the
efficacy of this picture~\citep{livneglasner91, livnearnett95,HoefKhok96,
hoeflichetal96, wigginsfalle97,wigginsetal98,garciasenzbravowoosley99}. 
These studies indicated that uncertainties in the initial conditions, 
however, were found to play a significant role on the explosion outcome.

A particular problem was that most models included a massive He layer on 
the white dwarf. Such a configuration leads to the unobserved result 
intermediate and heavy elements synthesized in the He detonation appear in 
the outer parts of the ejecta, which does not match obesrvation~\citep{HoefKhok96,
hoeflichetal96,finkhillebrandtroepke2007,simetal2010}.  \citet{bildstenetal2007}, 
however, found that fairly thin He layers could flash on sub-Chandrasekhar
mass white dwarfs, partially resolving the issue and encouraging
further research~\citep{simetal2012,brooksetal2015, shenetal2018,
glasneretal2018}.

The white dwarf merger progentor picture has two white dwarfs coming 
together and subsequently exploding~\citep{tutukovyungelson76,tutukovyungelson79,
webbink84,ibentutukov84}. This scenario provides an abundance of degenerate fuel, 
which may explain some bright events~\citep{scalzo:2010,Yuan:2010}. 
An early concern about this model came from studies finding that 
the as the stars merge, the more massive white dwarf can ignite near
its edge and a flame will burn inward, converting the C-O white dwarf into 
an O-Ne-Mg white dwarf~\citep{saionomoto1985,saionomoto2004}.  Further accretion 
leads to the collapse of the white dwarf
into a neutron star, a scenario termed ``accretion-induced collapse''
\citep{nomotokondo1991}.

Subsequent work allayed this concern~\citep{yoonetal2007,lorenaguilaretal2009,
Shenetal12, pakmoretal2012b} and motivated by the need to expain bright events
and the estimated population WD pairs, active research continues. Contemporary
research focuses on variations on the merger idea, including inspiraling pairs, 
collisions, violent mergers, and the ``core-degenerate'' model in which the merger 
takes place in a common envelope \citep{raskinetal2009,pakmoretal2011,kashi:2011,
pakmoretal2012a,Shenetal12,katzetal2016}.

\subsection{The Deflagraiton to Detonation Transition Mechanism Within 
the Single Degenerate Scenario}

Our single-degenerate models are within the DDT explosion paradigm \cite{1986SvAL,
Khokhlov1991Delayed-detonat,NiemWoos97,Niem99,belletal2004,fishjump2015}. 
In this case, the accretion of mass on the white dwarf compresses and 
heats the core, igniting carbon fusion and driving a period of convection
\citep{WoosWunsKuhl04,wunschwoosley2004,Kuhletal06,nonakaetal2012}.
At some point, the fusion rate becomes fast enough due to the rising
temperature that energy production exceeds convective cooling and
the deflagration phase begins in the core~\citep{Nomo84,WoosWunsKuhl04}.
This flame is unstable, and as the deflagration propagates toward the surface
of the WD, it is subject to the Rayleigh-Taylor instability that generates
turbulence and boosts buring.  
Burning proceeds as a deflagration for about one second, and then 
the flame transitions to a detonation~\citep{hoflich.khokhlov.ea:delayed}.

@@

The mechanism by which the deflagration transistions to a detonation
is imcompletely understood, but the development of fluid instabilities
at the deflagration front, particuarly the Rayleigh-Taylor 
instability, is an accepted part of the process.
Many ideas have been proposed, including the deflagration front
entering a regime of distributed burning and the net
rate becoming supersonic \citep{NiemWoos97}, the unstable
flame reaching a certian fractal dimension~\citep{woosley90},
and the Zel'dovich mechanism in which condtions are right for
a pocket of fuel to 
detonate~\citep{zeldovichetal1970,KhokOranWhee97,jacketal2014}.
We note that regardless of
the mechanism, the amount of expansion of the star prior to the bulk of burning
is critical to the composition of material synthesized in the explosion. Also,
early burning during the deflagtation phase is at high enough densites that
the effects of electron capture are significant and similarly influence the 
compositon the material synthesised in the explosion~\citep{hoeflichetal2004,
hoeflich2006,fesenetal2007,diamondetal2018}.
Our simulations assume that the transition occurs when the top of a 
rising, Rayleigh-Taylor unstable plume reaches a characteristic low
density~\citep{townsley.calder.ea:flame}. 
Variations on the choice of this density allowed earlier studies to explore
the effects of physics influencing the transition 
density~\citep{jacketal2010,Krueger2010On-Variations-o,kruegetal12}.

@@

In either {\color{red}ANY} case, the hybrid WD will experience a period of cooling, which will
make the core-mantle interface subject to convective 
instability \citep{brooksetal2017,schwabgaraud2018}. In the case of the 
hybrid WD accreting and approaching the Chandrasekhar mass, accretion will
heat the core and start C fusion, leading to a period of ``simmering" prior
to the explosion \citep{PiroBild08}. The upshot is that is likely to be considerable
mixing after the hybrid forms that may homegenize the compostion 
\citep{denissenkovetal2015,brooksetal2017,schwabgaraud2018}.

@@

From population synthesis, \citet{mengpods2014} found that these
%meng received may 5 accepted may 16
progenitors may substantially contribue to the population of \SNeIa (1-8\%) and have
relatively short delay times. They also suggested that these 
may produce part of the Iax class of events. \citet{Wangetal2014} also with population
synthesis studied the case
%received august 28 accepted sept 26
of a hybrid progenitor accreting from a nondegenerate He star and found
birth rates indicating that up to 18\% of SNe Ia may follow from this chanel
and very short delay times. \citet{Wangetal2014} also suggested that explosions 
from hybrid progenitors may provide an explanation for type Iax events.
\citet{mengpods2018}, from the common-envelope-wind model developed in 
\citep{mengpods2014}, propose that both Ia-CSM and Iax events 
are caused by the explosion of hybrid progenitors, with Ia-CSM occuring in systems with 
a massive common envelope and Iax events occuring in systems where most of the common envelope
has been lost.

@@

The methodology of our study follows the approach of \citet{willcoxetal2016}. 
We performed a suite of two-dimensional simulations of supernova explosions
from hybrid progenitors and compared the results to a suite of simulations
of explosions from traditional C/O progenitor models performed with the
same simulation code and from similar initial conditions. The simulations assumed
the deflagration to detonation transition explosion paradigm, and the
transition densities were the same in both suites. We briefly our methodoly
here, and refer to the reader to previous work for additional details.
In particular, we use the same simulation instrument, a modified version
of the Flash Code, as \citet{willcoxetal2016} and we refer the reader there 
for a description of modifications for C-O-Ne burning and verification tests
of the burning module.

@@

The simulations of supernova explosions presented here were performed
with a customized version of the \FLASH code, originally developed
at the University of Chicago.
Flash is a parallel, adaptive mesh, multi-physics simulation code
developed first for nuclear astrophysics applications and subsequently
for high-energy-density applications~\cite{Fryxetal00,calder.curtis.ea:high-performance,
calder.fryxell.ea:on,flash_pragmatic,flash_evolution}. 
Flash has been applied to a variety of astropysical problems by a host
of researchers, and the version we apply differs from other versions
principally in the modules describing thermonuclear burning via a 
flame capturing model.

The need for a model flame in simulating thermonuclear supernovae follows 
from the scales of the problem. At high densities, the width of a laminar 
nuclear flame is $< 1\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$ while the 
radius of the white dwarf $\sim 10^9\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$.
Even with adaptive mesh refinement, whole-star simulations cannot 
simultaneousely resolve the nuclear flame, so simulations require a model to 
describe the burning on unresovable scales.  The model we apply is a flame 
capturing scheme and thermally-activated burning module to describe
thermonuclear burning during both the deflagration and detonation phases, 
as well as routines to describe the evolution of the dynamic ash. 
This description of the burning was developed during the course of research in 
thermonuclear supernovae and has been presented in a series of 
papers \citep[See][and references therein]{chandlery}. For completeness, 
we briefly review the flame capturing scheme here.

For the deflagration phase, the flame capturing scheme propagates an 
artificially broadened flame with an advection-diffusion-reaction (ADR)
scheme~\cite{Khok95,VladWeirRyzh06} via evolution of
a reaction progress variable to describe the consumption of C
and additional variables to describe the evolution of intermediate-mass
elements into the stastical quasi-equilibruim of the 
Si-group~\cite{ifk1981,khok1981,khok1983} and then into iron-group elements
(IGEs) in full nuclear statistical equilibrium (NSE).

@@

The two-dimensional models in this study do not utilize a subgrid-scale
model for the turbulence-flame 
interaction~\citep[See][for examples]{Schmetal06a,Schmetal06b,jacketal2014},
The models only apply in three-dimensional
simulations beacuse two-dimensional hydrodynamics cannot
correctly describe turbulence. The simulations we present
use the mimimal enhancement based on the Rayleigh-Taylor
strength introduced by~\citet{townsley.calder.ea:flame}. The
assumption here is that the burning self-regulate on resolved 
scales so that results are insensitive to the detailed treatment 
of the interaction with turbulence, and previous experience
indicates this assumption is reasonable for comparisons
like the one presented in this work~\citet{townsley.calder.ea:flame,
willcoxetal2016}.

@@

The hybrid white dwarf model that served as the initial conditions 
for the simulations of supernova explosions presented in this work
was constructed with the one-dimensional stellar evolution code 
MESA~\citep{mesa1,mesa2,mesa3,mesa3e}. The evolution of the model 
included thermohalene mixing during the cooling phase and thus
had a lower C fraction than previous hybrid models~\citep{brooksetal2017}.
The hybrid model was constructed to have conditions as similar a possible
to a ``classic" C/O model and the two shared the same central temperature
and density.  Figure~\ref{fig:init_conds} shows the density and temperature
profiles of the two initial one-dimensional models. 
\begin{figure}
\includegraphics[width=\columnwidth]{figures/initial_conds.png}
\caption{\label{fig:init_conds}
Temperature, density, and  composition profiles of the one-dimensional hybrid 
progenitor WD. The central temperature and density were the same as
the tradtional C/O progenitor model. 
The peak temperature of the hybrid model is at the center of the star, 
so both the hybrid and traditional models could share the same 
	central ignition initial conditions.
}
\end{figure}
In both models, and unlike the hybrid models of~\citet{willcoxetal2016}, the central
temperature was the highest at the center of the star and accordingly the simulations 
began with central ignitions.

@@

The process of mapping the \MESA\ models began by converting 
the \MESA\ model to a uniform grid of $4$~km resolution. \MESA\ 
quantities in zones with spacing less than $4$~km 
were mass-weighted averaged and a quadratic interpolation
was used to calculate quantities for the case of \MESA\ zones having
a spacing greater than $4$~km. At this point, 
abundances of nuclides from the \MESA\ models were aggregated
into the abundances tracked by the flame capturing scheme in
\FLASH.
The most abundant isotopes in the model were \C{12}, \Ox{16},
and \Ne{20}, which are symmetric (the number of neutrons 
equals the number of protons). Similarly to~\citet{willcoxetal2016}, 
other neutron-rich isotopes in the initial model were combined into 
\Ne{22}, which serves as a proxy for metallicty. The aggregation
account for the $Y_e$ of the full set of nuclides, \Ne{20} 
and \Ox{16} were constrained to be in the same ratio in both 
sets of abundances. \figref{fig:init_conds} shows the initial
profile of the one-dimensional models.

The original \MESA\ model was in equilibrium, but to ensure
the uniformly gridded model was in hydrostatic equilibrium we 
constructed the appropriate pressure profile. This was done by
integrating for the pressure in each zone from the central point up, 
accounting for the local acceleration of gravity, temperature,
composition, and the mass below and enclosed by the zone. As
with \citet{willcoxetal2016} we used
EOS routine from 
\CASTRO\ ~\citep{timmes.swesty:accuracy,castro1}.\footnote{The 
Helmholtz EOS table used in \CASTRO\ is in the public BoxLib 
Microphysics repository at
  \url{https://github.com/BoxLib-Codes/Microphysics.git} (commit 
  hash $\mathrm{45ed859b6c1dc80d831d93f9728986d6ad6e1ddc}$).} As 
 with the models in \citet{willcoxetal2016},this procedure produced 
 a structure that was stable in \FLASH, with fluctuations in central
density less than $3\%$, for at least $5$ seconds with no energy
deposition.

@@

The simulations performed for this study consisted to a suite of
{\color{red} 35 CHECK THIS! 40?} two-dimensional simulations of 
thermonuclear supernova
explosions from hybrid C-O-Ne progenitors in the DDT explosion paradigm. These 
were compared to a suite of supernova simulations from traditional
C-O progenitors. In both suites, the simulation
begins with a progenitor model mapped to the two-dimensional 
\FLASH\ grid. The burning is initiated with a ``matchhead,'' a region 
in the white dwarf's center that is fully burned to nuclear statistical
equilibrium (NSE). 
This initially burned regeion ignites a deflagration, a subsonic
flame, and because the mathhead was perturbed it is unstable to
the Rayleigh-Taylor instability with the result that buoyant plumes 
rise. The star is partially consumed during this deflagration phase,
and the star responds by expanding.
When a plume reaches a specified density,
a detonation is initiated, and the simulation continues until 
the expanding star reaches low densities at which point burning
effectively ends. The paragraphs in this section provide details
of the implementation of this method.

>>>>

\begin{abstract}
The existence of ``hybrid" white dwarfs, made of a C/O
core within a O/Ne shell is accepted, and studies indicate that these are 
viable progenitors for thermonuclear (type Ia) supernovae.
Recent work found that the C/O core is mixed with the surrounding O/Ne
as the white dwarf cools prior to accretion, which results in lower central 
C fractions in the massive progenitor than previously assumed. To further 
investigate the efficacy of hybrid white dwarfs as progenitors of 
thermonuclear supernovae, we performed simulations of thermonuclear 
supernovae from a new series of hybrid progenitors that include the 
effects of cooling. The progenitor white dwarf model
was constructed with the one-dimensional stellar evolution code MESA and
represented a star evolved through the phase of unstable interior mixing 
followed by accretion until it reached conditions for the ignition of 
carbon burning. This MESA model was then mapped to a two-dimensional initial 
condition and an explosion simulated with FLASH. For comparison, a similar
simulation of an explosion was performed from a traditional C/O progenitor
white dwarf. By comparing the yields of the explosions, we found that as
with earlier studies, these hybrid models are viable supernova progenitors.
We contrast explosions from these models with explosions 
from previous hybrid models 
{\color{red} without the mixing during the white dwarf cooling}
and traditional C/O progenitors.
\end{abstract}

@@

Supernovae are classified observationally 
by their light curves and spectra~\citep{minkowski41,bertola64,porterfilippenko87,
wheelerharkness1990conf,Fili97}, with the type Ia designation following from
the absence of H in the spectrum and the presence of a specific Si 
line~\citep{filippenko:optical,hillebrandt.niemeyer:type}. These events
have been associated with C burning under degenerate conditions
for some time~\citep{hoylefowler60,arnett.truran.ea:nucleosynthesis},
but discerning the setting(s) of these events is proving difficult
and remains the subject of active research. At present there are three
widely-accepted scenarios: the {\em single degenerate} scenario,
the {\em double detonation} or {\em sub-Chandrasekhar} scenario, 
the {\em white dwarf merger} or {\em double degenerate} scenario.
We briefly describe these in the subsection that follows.
Also see~\citet{hillebrandt.niemeyer:type,howell2011,hillebrandtetal2013,calderetal2013,roepkesim2018}
for additional discussion.

@@

The single degenerate picture posits a white dwarf gaining mass
from a main-sequence companion. The process relies on a long
period of accretion combined with either steady burning or a 
series of nova explosions to allow the WD to gain at least
$\sim 0.4$ \Msun\ as needed for it to approach the
limiting Chandrasekhar mass~\citep{starrfieldetal2012}. As it approaches
the Chandrasekhar limit, conditions in the compressed core are right
to ignite the thermonuclear burning that will incinerate the star.

Within this progenitor setting, many mechanisms for the explosion have 
been studied with the result being that the outcome is sensitive to
the ignition conditions and the nature of the thermonuclear burning.
Studies indicate that neither a pure deflagration (subsonic flame front)
nor a pure detonation (supersonic burning front) produce ejecta consistent
with observations \citep{arnett69,roepkeetal07}.  Instead, the models 
that best reproduce the observed stratified ejecta are those in which 
the burning begins as a subsonic deflagration, which allows the star 
to react and expand thus lowering the density, and is then followed by a 
supersonic detonation that incinerates the 
star~\citep{Nomo84,Khokhlov1991Delayed-detonat,HoefKhok96,GameKhokOran05}.
The ``classic'' delayed detonation model is that of 
\citet{Khokhlov1991Delayed-detonat} \citep[See also][]{hoflich.khokhlov.ea:delayed,GameKhokOran05},
and many variations have been 
explored \citep[and references therein]{hillebrandtetal2013,calderetal2013}.
The approach we employ for this study is a variation, the 
deflagration-to-detonation transition (DDT), in which the transition
occurs at the top of a rising Rayleigh-Taylor~\citep{taylor+50,chandra+81} 
unstable plume of hot, burned 
material when it reaches a threshold density.
We describe this approach and our methodology in detail below.

The double detonation picture is a variation of the single degenerate picture.
In this case, a white dwarf accretes material from a companion, but rather
than gaining enough mass to approach the Chandrasekhar limit, a detonation
occurs in the accreted layer that subsequently triggers a detonation
in the underlying white dwarf \citet{woosleyweavertaam80,taam80a,taam80b,
nomoto80,nomoto82b}. Early studies indicated that a detonation in an 
accreted He
layer could produce an inwardly propagating
shock that would ignite a detonation in the C-O core and found
that this scenario could work for a wide range of white dwarf
masses, not just the near-Chandrasekhar case \citep{livne90}, hence the
moniker ``sub-Chandrasekhar'' \citep{ww94}. Multidimensional supported the
efficacy of this picture~\citep{livneglasner91, livnearnett95,HoefKhok96,
hoeflichetal96, wigginsfalle97,wigginsetal98,garciasenzbravowoosley99}. 
These studies indicated that uncertainties in the initial conditions, 
however, were found to play a significant role on the explosion outcome.

A particular problem was that most models included a massive He layer on 
the white dwarf. Such a configuration leads to the unobserved result 
intermediate and heavy elements synthesized in the He detonation appear in 
the outer parts of the ejecta, which does not match observation~\citep{HoefKhok96,
hoeflichetal96,finkhillebrandtroepke2007,simetal2010}.  \citet{bildstenetal2007}, 
however, found that fairly thin He layers could flash on sub-Chandrasekhar
mass white dwarfs, partially resolving the issue and encouraging
further research~\citep{simetal2012,brooksetal2015, shenetal2018,
glasneretal2018}.

The white dwarf merger progenitor picture has two white dwarfs coming 
together and subsequently exploding~\citep{tutukovyungelson76,tutukovyungelson79,
webbink84,ibentutukov84}. This scenario provides an abundance of degenerate fuel, 
which may explain some bright events~\citep{scalzo:2010,Yuan:2010}. 
An early concern about this model came from studies finding that 
the as the stars merge, the more massive white dwarf can ignite near
its edge and a flame will burn inward, converting the C-O white dwarf into 
an O-Ne-Mg white dwarf~\citep{saionomoto1985,saionomoto2004}.  Further accretion 
leads to the collapse of the white dwarf
into a neutron star, a scenario termed ``accretion-induced collapse''
\citep{nomotokondo1991}.

Subsequent work allayed this concern~\citep{yoonetal2007,lorenaguilaretal2009,
Shenetal12, pakmoretal2012b} and motivated by the need to explain bright events
and the estimated population WD pairs, active research continues. Contemporary
research focuses on variations on the merger idea, including inspiraling pairs, 
collisions, violent mergers, and the ``core-degenerate'' model in which the merger 
takes place in a common envelope \citep{raskinetal2009,pakmoretal2011,kashi:2011,
pakmoretal2012a,Shenetal12,katzetal2016}.

\subsection{The Deflagration to Detonation Transition Mechanism Within 
the Single Degenerate Scenario}

Our single-degenerate models are within the DDT explosion paradigm \cite{1986SvAL,
Khokhlov1991Delayed-detonat,NiemWoos97,Niem99,belletal2004,fishjump2015}. 
In this case, the accretion of mass on the white dwarf compresses and 
heats the core, igniting carbon fusion and driving a period of convection
\citep{WoosWunsKuhl04,wunschwoosley2004,Kuhletal06,nonakaetal2012}.
At some point, the fusion rate becomes fast enough due to the rising
temperature that energy production exceeds convective cooling and
the deflagration phase begins in the core~\citep{Nomo84,WoosWunsKuhl04}.
This flame is unstable, and as the deflagration propagates toward the surface
of the WD, it is subject to the Rayleigh-Taylor instability that generates
turbulence and boosts burning.  
Burning proceeds as a deflagration for about one second, and then 
the flame transitions to a detonation~\citep{hoflich.khokhlov.ea:delayed}.

@@

The mechanism by which the deflagration transitions to a detonation
is incompletely understood, but the development of fluid instabilities
at the deflagration front, particularly the Rayleigh-Taylor 
instability, is an accepted part of the process.
Many ideas have been proposed, including the deflagration front
entering a regime of distributed burning and the net
rate becoming supersonic \citep{NiemWoos97}, the unstable
flame reaching a certain fractal dimension~\citep{woosley90},
and the Zel'dovich mechanism in which conditions are right for
a pocket of fuel to 
detonate~\citep{zeldovichetal1970,KhokOranWhee97,jacketal2014}.
We note that regardless of
the mechanism, the amount of expansion of the star prior to the bulk of burning
is critical to the composition of material synthesized in the explosion. Also,
early burning during the deflagration phase is at high enough densities that
the effects of electron capture are significant and similarly influence the 
composition the material synthesized in the explosion~\citep{hoeflichetal2004,
hoeflich2006,fesenetal2007,diamondetal2018}.
Our simulations assume that the transition occurs when the top of a 
rising, Rayleigh-Taylor unstable plume reaches a characteristic low
density~\citep{townsley.calder.ea:flame}. 
Variations on the choice of this density allowed earlier studies to explore
the effects of physics influencing the transition 
density~\citep{jacketal2010,Krueger2010On-Variations-o,kruegetal12}.

@@

In either {\color{red}ANY} case, the hybrid WD will experience a period of cooling, which will
make the core-mantle interface subject to convective 
instability \citep{brooksetal2017,schwabgaraud2018}. In the case of the 
hybrid WD accreting and approaching the Chandrasekhar mass, accretion will
heat the core and start C fusion, leading to a period of ``simmering" prior
to the explosion \citep{PiroBild08}. The upshot is that is likely to be considerable
mixing after the hybrid forms that may homogenize the composition 
\citep{denissenkovetal2015,brooksetal2017,schwabgaraud2018}.

@@

From population synthesis, \citet{mengpods2014} found that these
%meng received may 5 accepted may 16
progenitors may substantially contribute to the population of \SNeIa (1-8\%) and have
relatively short delay times. They also suggested that these 
may produce part of the Iax class of events. \citet{Wangetal2014} also with population
synthesis studied the case
%received august 28 accepted sept 26
of a hybrid progenitor accreting from a non-degenerate He star and found
birth rates indicating that up to 18\% of SNe Ia may follow from this channel
and very short delay times. \citet{Wangetal2014} also suggested that explosions 
from hybrid progenitors may provide an explanation for type Iax events.
\citet{mengpods2018}, from the common-envelope-wind model developed in 
\citep{mengpods2014}, propose that both Ia-CSM and Iax events 
are caused by the explosion of hybrid progenitors, with Ia-CSM occurring in systems with 
a massive common envelope and Iax events occurring in systems where most of the common envelope
has been lost.

@@

The methodology of our study follows the approach of \citet{willcoxetal2016}. 
We performed a suite of two-dimensional simulations of supernova explosions
from hybrid progenitors and compared the results to a suite of simulations
of explosions from traditional C/O progenitor models performed with the
same simulation code and from similar initial conditions. The simulations assumed
the deflagration to detonation transition explosion paradigm, and the
transition densities were the same in both suites. We briefly our methodology
here, and refer to the reader to previous work for additional details.
In particular, we use the same simulation instrument, a modified version
of the Flash Code, as \citet{willcoxetal2016} and we refer the reader there 
for a description of modifications for C-O-Ne burning and verification tests
of the burning module.

@@

The simulations of supernova explosions presented here were performed
with a customized version of the \FLASH code, originally developed
at the University of Chicago.
Flash is a parallel, adaptive mesh, multi-physics simulation code
developed first for nuclear astrophysics applications and subsequently
for high-energy-density applications~\cite{Fryxetal00,calder.curtis.ea:high-performance,
calder.fryxell.ea:on,flash_pragmatic,flash_evolution}. 
Flash has been applied to a variety of astrophysical problems by a host
of researchers, and the version we apply differs from other versions
principally in the modules describing thermonuclear burning via a 
flame capturing model.

The need for a model flame in simulating thermonuclear supernovae follows 
from the scales of the problem. At high densities, the width of a laminar 
nuclear flame is $< 1\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$ while the 
radius of the white dwarf $\sim 10^9\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$.
Even with adaptive mesh refinement, whole-star simulations cannot 
simultaneously resolve the nuclear flame, so simulations require a model to 
describe the burning on unresovable scales.  The model we apply is a flame 
capturing scheme and thermally-activated burning module to describe
thermonuclear burning during both the deflagration and detonation phases, 
as well as routines to describe the evolution of the dynamic ash. 
This description of the burning was developed during the course of research in 
thermonuclear supernovae and has been presented in a series of 
papers \citep[See][and references therein]{chandlery}. For completeness, 
we briefly review the flame capturing scheme here.

For the deflagration phase, the flame capturing scheme propagates an 
artificially broadened flame with an advection-diffusion-reaction (ADR)
scheme~\cite{Khok95,VladWeirRyzh06} via evolution of
a reaction progress variable to describe the consumption of C
and additional variables to describe the evolution of intermediate-mass
elements into the statistical quasi-equilibrium of the 
Si-group~\cite{ifk1981,khok1981,khok1983} and then into iron-group elements
(IGEs) in full nuclear statistical equilibrium (NSE).

@@

The two-dimensional models in this study do not utilize a sub-grid-scale
model for the turbulence-flame 
interaction~\citep[See][for examples]{Schmetal06a,Schmetal06b,jacketal2014},
The models only apply in three-dimensional
simulations because two-dimensional hydrodynamics cannot
correctly describe turbulence. The simulations we present
use the minimal enhancement based on the Rayleigh-Taylor
strength introduced by~\citet{townsley.calder.ea:flame}. The
assumption here is that the burning self-regulate on resolved 
scales so that results are insensitive to the detailed treatment 
of the interaction with turbulence, and previous experience
indicates this assumption is reasonable for comparisons
like the one presented in this work~\citet{townsley.calder.ea:flame,
willcoxetal2016}.

@@

The hybrid white dwarf model that served as the initial conditions 
for the simulations of supernova explosions presented in this work
was constructed with the one-dimensional stellar evolution code 
MESA~\citep{mesa1,mesa2,mesa3,mesa3e}. The evolution of the model 
included thermohalene mixing during the cooling phase and thus
had a lower C fraction than previous hybrid models~\citep{brooksetal2017}.
The hybrid model was constructed to have conditions as similar a possible
to a ``classic" C/O model and the two shared the same central temperature
and density.  Figure~\ref{fig:init_conds} shows the density and temperature
profiles of the two initial one-dimensional models. 
\begin{figure}
\includegraphics[width=\columnwidth]{figures/initial_conds.png}
\caption{\label{fig:init_conds}
Temperature, density, and  composition profiles of the one-dimensional hybrid 
progenitor WD. The central temperature and density were the same as
the traditional C/O progenitor model. 
The peak temperature of the hybrid model is at the center of the star, 
so both the hybrid and traditional models could share the same 
	central ignition initial conditions.
}
\end{figure}
In both models, and unlike the hybrid models of~\citet{willcoxetal2016}, the central
temperature was the highest at the center of the star and accordingly the simulations 
began with central ignitions.

@@

The process of mapping the \MESA\ models began by converting 
the \MESA\ model to a uniform grid of $4$~km resolution. \MESA\ 
quantities in zones with spacing less than $4$~km 
were mass-weighted averaged and a quadratic interpolation
was used to calculate quantities for the case of \MESA\ zones having
a spacing greater than $4$~km. At this point, 
abundances of nuclides from the \MESA\ models were aggregated
into the abundances tracked by the flame capturing scheme in
\FLASH.
The most abundant isotopes in the model were \C{12}, \Ox{16},
and \Ne{20}, which are symmetric (the number of neutrons 
equals the number of protons). Similarly to~\citet{willcoxetal2016}, 
other neutron-rich isotopes in the initial model were combined into 
\Ne{22}, which serves as a proxy for metallicity. The aggregation
account for the $Y_e$ of the full set of nuclides, \Ne{20} 
and \Ox{16} were constrained to be in the same ratio in both 
sets of abundances. \figref{fig:init_conds} shows the initial
profile of the one-dimensional models.

The original \MESA\ model was in equilibrium, but to ensure
the uniformly-gridded model was in hydrostatic equilibrium we 
constructed the appropriate pressure profile. This was done by
integrating for the pressure in each zone from the central point up, 
accounting for the local acceleration of gravity, temperature,
composition, and the mass below and enclosed by the zone. As
with \citet{willcoxetal2016} we used
EOS routine from 
\CASTRO\ ~\citep{timmes.swesty:accuracy,castro1}.\footnote{The 
Helmholtz EOS table used in \CASTRO\ is in the public BoxLib 
Microphysics repository at
  \url{https://github.com/BoxLib-Codes/Microphysics.git} (commit 
  hash $\mathrm{45ed859b6c1dc80d831d93f9728986d6ad6e1ddc}$).} As 
 with the models in \citet{willcoxetal2016},this procedure produced 
 a structure that was stable in \FLASH, with fluctuations in central
density less than $3\%$, for at least $5$ seconds with no energy
deposition.

@@

The simulations performed for this study consisted to a suite of
{\color{red} 35 CHECK THIS! 40?} two-dimensional simulations of 
thermonuclear supernova
explosions from hybrid C-O-Ne progenitors in the DDT explosion paradigm. These 
were compared to a suite of supernova simulations from traditional
C-O progenitors. In both suites, the simulation
begins with a progenitor model mapped to the two-dimensional 
\FLASH\ grid. The burning is initiated with a ``match head,'' a region 
in the white dwarf's center that is fully burned to nuclear statistical
equilibrium (NSE). 
This initially burned region ignites a deflagration, a subsonic
flame, and because the match head was perturbed it is unstable to
the Rayleigh-Taylor instability with the result that buoyant plumes 
rise. The star is partially consumed during this deflagration phase,
and the star responds by expanding.
When a plume reaches a specified density,
a detonation is initiated, and the simulation continues until 
the expanding star reaches low densities at which point burning
effectively ends. The paragraphs in this section provide details
of the implementation of this method.
